\documentclass[11pt]{article}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}


\title{DS 542 - Spring 2026 - Homework 1}
\date{}

\begin{document}
%\vspace*{-20pt}
\maketitle

\noindent
Feel free to write your answers by hand.
No need for LaTeX; just make sure it is legible.

\begin{enumerate}

\item

  Equation 2.5 in the textbook (Understanding Deep Learning) gives the equation for linear regression with one input variable using the least-squares loss.
  \begin{eqnarray*}
    L[\phi] & = & \sum_{i=1}^{I} \left( f[x_i,\phi] - y_i \right)^2 \\
    & = & \sum_{i=1}^{I} \left( \phi_0 + \phi_1 x_i - y_i \right)^2
  \end{eqnarray*}

  \begin{enumerate}

  \item

    \textbf{(2 points)} Derive the formula for \( \partial L / \partial \phi_0 \).

    \vspace{\stretch{1}}

  \item
    \textbf{(2 points)} Derive the formula for \( \partial L / \partial \phi_1 \).

    \vspace{\stretch{1}}
  \end{enumerate}

  \newpage
  
\item

  Answer the following questions about the following equation.
  \begin{eqnarray*}
    \begin{bmatrix} 1 & 2 \\ 4 & 3 \\ \end{bmatrix}
    \begin{bmatrix} 8 & 7 \\ 5 & 6 \\ \end{bmatrix}
    & = &
    \begin{bmatrix} x_{0,0} & x_{0,1} \\ x_{1,0} & x_{1,1} \\ \end{bmatrix}
  \end{eqnarray*}

  \begin{enumerate}
    
  \item
    
    \textbf{(2 points)} What is the mathematical operation on the lefthand side of the equation?

    \vspace{2cm}

  \item
    \textbf{(2 points)} Write out the formula to calculate $x_{1,0}$.

    \vspace{2cm}
  \end{enumerate}

\item

  \textbf{(2 points)} Answer the following questions about ordinary least squares linear regression, where $y_i$ is the target value and $\hat{y}_i$ is the predicted value.
  What is the equation for the mean residual on the training data?

  \vspace{2cm}

\item

  Answer the following questions about basic probability.

  \begin{enumerate}

  \item
    \textbf{(2 points)} Suppose \(A\) and \(B\) are events with \(P(A)=0.4\), \(P(B)=0.5\), and \(P(A\cap B)=0.2\).
    Compute \(P(A \mid B)\) and \(P(A\cup B)\).

    \vspace{2cm}

  \item
    \textbf{(2 points)} Let \(X\) be a Bernoulli random variable with \(P(X=1)=p\).
    What are \(\mathbb{E}[X]\) and \(\mathrm{Var}(X)\)?

    \vspace{2cm}

  \end{enumerate}

  
\end{enumerate}

\end{document}  
